### 2021-04-28 submission/manuscript.tex
Figure legend edits and additions

\textbf{5-day laxative treatment leads to prolonged \emph{C. difficile}
colonization in mice.} We compared mice treated with the osmotic
laxative PEG 3350 to our standard 10 mg/kg clindamycin treatment, which
temporarily renders the mice susceptible to \emph{C. difficile}, with
mice typically clearing \emph{C. difficile} within 10 days
post-infection (9, 26). All PEG-treated mice were administered a 15\%
PEG solution in the drinking water for 5-days, one group was also
treated with clindamycin, and one group was allowed to recover for 10
days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss
in all 3 groups of PEG-treated mice, with the greatest change in weight
observed on the fifth day of PEG treatment and the mice recovered most
of the weight five days after treatment (Fig. 1B). After either PEG,
clindamycin, or PEG and clindamycin treatment all mice were challenged
with 10\textsuperscript{3} \emph{C. difficile} 630 spores (Fig. 1A). All
treatments rendered mice susceptible to \emph{C. difficile}
colonization. However, PEG-treated mice remained colonized with \emph{C.
difficile} at a high level through thirty days post-infection (Fig. 1C).
In contrast, the clindamycin-treated mice cleared \emph{C. difficile}
within ten days post-infection. Surprisingly, PEG-treated mice were
still susceptible to \emph{C. difficile} infection after 10-days of
recovery from treatment, although \emph{C. difficile} was not detectable
in most of the group in the initial five days post-infection (Fig. 1C).
From 8 days post-infection onward, the median \emph{C. difficile} CFU
stabilized for the 10-day recovery group of PEG-treated mice and
remained high through 30 days post-infection (Fig. 1C). Thus, osmotic
laxative treatment alone was sufficient to render mice susceptible to
prolonged \emph{C. difficile} colonization and PEG-treated mice remained
susceptible up to ten days post-treatment.

@@

\textbf{5-day laxative treatment does not promote more severe CDIs
despite altering the mucosal microbiota.} Given the findings from a
previous study that demonstrated PEG treatment disrupts the mucus layer
and alters the immune response in mice (16), we decided to examine the
impact of PEG treatment on the mucosal microbiota and CDI severity. To
evaluate the mucosal microbiota, we sequenced snips of tissue collected
from the cecum, proximal colon, and distal colon. Similar to what was
observed with the stool samples, alpha diversity was lower in the
PEG-treated mice compared to clindamycin treatment (Fig. 3A). Although
alpha diversity continued to increase over time based on the communities
from PEG-treated mice collected at 20 and 30 days post-infection (Fig.
3A, Data Set S1, Sheet X). Group, time point, and their interactions
with other variables (cage, experiment number, and sample type)
explained the majority of the variation observed in mucosal communities
(PERMANOVA combined R\textsuperscript{2} = 0.83, \emph{P} \textless{}
0.05, Fig. 3B, Data Set S1, Sheet X). We saw the greatest difference in
the mucosal microbiota between treatment groups at 6 days post-infection
with 16 genera that were significantly different in all three of the
tissue types we collected (cecum, proximal colon, and distal colon; Fig.
S3A, Data Set S1, Sheet X). However, at 30 days post-infection only
\emph{Bacteroides}, \emph{Clostridiales}, \emph{Firmicutes}, and
\emph{Ruminococcaceae} were different between treatment groups and only
in the cecum tissues (Fig. 3C, Fig. 2E, Data Set S1, Sheet X).
Interestingly, one of the bacteria that was different between treatment
groups at 6 days post-infection was \emph{Peptostreptococcaceae} (the
genera with an OTU sequence that matches \emph{C. difficile}).
\emph{Peptostreptococcaceae} was primarily only present in the 5-day PEG
treatment group of mice and decreased in the proximal and distal colon
tissues over time (Fig. S2B). Thus, PEG treatment had a significant
impact on the mucosal microbiota and we detected \emph{C. difficile}
sequences in the cecum, proximal colon, and distal colon tissue
communities.

@@

Previous work examining the impact of PEG on the murine microbiota found
that PEG treatment resulted in the permanent loss of S24-7, also known
as \emph{Muribaculum intestinale} (16). We decided to check our
\emph{Porphyromonadaceae} OTUs because \emph{Muribaculum intestinale} is
classified as \emph{Porphyromonadaceae} by the Ribosomal Database
Project (RDP) database (31) and \emph{Porphyromonadaceae} was a top
feature in the random forest model predicting prolonged \emph{C.
difficile} colonization. We identified 4 OTUs that had at least 92\%
identity to \emph{Muribaculum intestinale} and examined their abundance
in mice that either cleared or were still colonized with \emph{C.
difficile} at 10 days post-infection. While all of the OTUs, were
decreased by PEG and clindamycin treatment, there was some recovery in
the mice that cleared (Fig. S4A). We also examined other
\emph{Porphyromonadaceae} and \emph{Lachnospiraceae} OTUs since these
were the 2 genera that were important to our classification model and
contained multiple OTUs that were different at 5 days post-infection
between mice that either cleared or remained colonized with \emph{C.
difficile} by 10 days post-infection (Data Set S1, sheet X). While
individual \emph{Porphyromonadaceae} and \emph{Lachnospiraceae} OTUs
tended to be more abundant in the mice that clear \emph{C. difficile}
relative to the mice that exhibit prolonged colonization (Fig. S4B-C),
there is no single OTU that fits the pattern we observed at the genus
level (Fig. 7C), suggesting multiple \emph{Porphyromonadaceae} and
\emph{Lachnospiraceae} OTUs influenced \emph{C. difficile} clearance.
Overall, our results suggest that specific bacterial community
differences explain the prolonged \emph{C. difficile} colonization we
observed in 5-day PEG and post-CDI 1-day PEG treated mice.

@@

\textbf{Figure 1. 5-day PEG treatment prolongs susceptibility and mice
become persistently colonized with \emph{C. difficile}.} A. Setup of the
experimental timeline for experiments with 5-day PEG treated mice.
Clindamycin was administered at 10 mg/kg by intraperitoneal injection.
15\% PEG 3350 was administered in the drinking water for five days. B.
Weight change from baseline weight in groups after treatment with PEG
and/or clindamycin, followed by \emph{C. difficile} challenge. C.
\emph{C. difficile} CFU/gram stool measured over time (N = 16-59 mice
per time point) via serial dilutions. The black line represents the
limit of detection for the first serial dilution. CFU quantification
data was not available for each mouse due to stool sampling difficulties
(particularly the day the mice came off of the PEG treatment) or early
deaths. For B-C, lines represent the median for each treatment group and
circles represent samples from individual mice. Asterisks indicate time
points where the weight change or CFU/g was significantly different
between groups by the Kruskal-Wallis test with Benjamini-Hochberg
correction for testing multiple time points. The data presented are from
a total of 5 separate experiments. \newpage

\includegraphics{figure_2.pdf} \textbf{Figure 2. 5-day PEG treatment
disrupts the stool microbiota for a longer amount of time compared to
clindamycin-treated mice.} A. PCoA of Bray-Curtis distances from stool
samples collected throughout the experiment. \newpage

\includegraphics{figure_3.pdf} \textbf{Figure 3. 5-day PEG treatment
does not result in more severe CDIs, although mucosal microbiota is
altered.} A. \newpage

\includegraphics{figure_4.pdf} \textbf{Figure 4. 1-day PEG treatment
renders mice susceptible to transient \emph{C. difficile} colonization.}
A. Setup of the experimental timeline for the 1-day PEG treated subset
of mice. B. CFU/gram stool measured over time (N = 12-18 mice per time
point) via several dilutions. The black dotted line represents the limit
of detection for the first serial dilution. Asterisks indicate time
points where the CFU/gram was significantly different between groups
using the Kruskall-Wallis test with a Benjamini-Hochberg correction for
multiple time points. C. Principle Coordinate Analysis plot of the
groups over time with the alpha representating the same time scale as in
panel D (day: R\textsuperscript{2} = 0.43; group: R\textsuperscript{2} =
0.19). D. Shannon diversity Index of the groups over time. Only days
with samples from all groups are shown. E. Line plots of relative
percent abundance of selected genera over time. Only days with samples
from all groups shown. The gray line represents the limit of detection.

\includegraphics{figure_5.pdf} \textbf{Figure 5. 1-day PEG treatment
post \emph{C. difficile} challenge prolongs colonization regardless of
whether an FMT is also administered.} All mice were administered 10
mg/kg clindamycin intraperitoneally (IP) 1 day before challenge with
\emph{C. difficile} 630 spores on day 0. 15\% PEG 3350 was administered
in the drinking water for 1 day. Mice feces were collected daily through
the end of the experiment (30 days post-infection for Clind. + 1-day PEG
group, 15 days post-infection for all others). CFU quantification data
was not available for each mouse due to stool sampling difficulties
(particularly the day the mice came off of the PEG treatment) or early
deaths. B. CFU/g of \emph{C. difficile} stool measured over time via
serial dilutions. C. Median and individual Shannon Diversity measured
over time. D. Median and individual richness measured over time. B-D.
Asterisks indicate time points with significant differences between
groups using Kruskall-Wallis tests with a Benjamini-Hochberg correction
for testing multiple times points. Black line represents the limit of
detection for the first serial dilution. Background shading indicates
PEG treatment period for applicable groups. Opacity of points reflects
time point. The data presented are from a total of *** separate
experiments. \newpage

\includegraphics{figure_6.pdf} \textbf{Figure 6. For 1-day PEG treatment
post \emph{C. difficile} challenge mice that also receive an FMT only
some bacterial genera were restored.} A. PCoA of all groups + FMT
plotted over time. Opacity of points reflects time point. B. Relative
abundance of genera significantly different over multiple time points
post FMT or PBS treatment, plotted over time. Limit of detection is
.1\%. C. Relative abundances of top 6 significant genera ranked by
number of days significant, plotted over time. Limit of detection is
.1\%.

@@

\includegraphics{figure_7.pdf} \textbf{Figure 7. Specific microbiota
features associated with prolonged \emph{C. difficile} colonization in
PEG treated mice.} A. \newpage

\includegraphics{figure_8.pdf} \textbf{Figure 8. Schematic summarizing
findings.} A. \newpage

\includegraphics{figure_S1.pdf} \textbf{Figure S1. 5-day PEG treatment
plus 10-day recovery mice microbiota dynamics post-infection.} A.
\newpage

\includegraphics{figure_S2.pdf} \textbf{Figure S2. Impact of PEG
treatment on the mucosal microbiota is greatest 6 days post-infection}\\
\newpage

@@

\includegraphics{figure_S4.pdf} \textbf{Figure S4. Specific OTUs
associated with clearance that are mostly absent in mice with prolonged
\emph{C. difficile} colonization. Ex. \emph{Muribaculum intestinale}.}
A. \newpage

>>>>

\textbf{5-day laxative treatment leads to prolonged \emph{C. difficile}
colonization in mice.} We compared mice treated with the osmotic
laxative PEG 3350 to our standard 10 mg/kg clindamycin treatment, which
temporarily renders the mice susceptible to \emph{C. difficile}, with
mice typically clearing \emph{C. difficile} within 10 days
post-infection (9, 26). All PEG-treated mice were administered a 15\%
PEG solution in the drinking water for 5-days, one group was also
treated with clindamycin, and one group was allowed to recover for 10
days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss
in all 3 groups of PEG-treated mice, with the greatest change in weight
observed on the fifth day of PEG treatment and the mice recovered most
of the weight five days after treatment (Fig. 1B). After either PEG,
clindamycin, or PEG and clindamycin treatment all mice were challenged
with 10\textsuperscript{5} \emph{C. difficile} 630 spores (Fig. 1A). All
treatments rendered mice susceptible to \emph{C. difficile}
colonization. However, PEG-treated mice remained colonized with \emph{C.
difficile} at a high level through thirty days post-infection (Fig. 1C).
In contrast, the clindamycin-treated mice cleared \emph{C. difficile}
within ten days post-infection. Surprisingly, PEG-treated mice were
still susceptible to \emph{C. difficile} infection after 10-days of
recovery from treatment, although \emph{C. difficile} was not detectable
in most of the group in the initial five days post-infection (Fig. 1C).
From 8 days post-infection onward, the median \emph{C. difficile} CFU
stabilized for the 10-day recovery group of PEG-treated mice and
remained high through 30 days post-infection (Fig. 1C). Thus, osmotic
laxative treatment alone was sufficient to render mice susceptible to
prolonged \emph{C. difficile} colonization and PEG-treated mice remained
susceptible up to ten days post-treatment.

@@

\textbf{5-day laxative treatment does not promote more severe CDIs
despite altering the mucosal microbiota.} Given the findings from a
previous study that demonstrated PEG treatment disrupts the mucus layer
and alters the immune response in mice (16), we decided to examine the
impact of PEG treatment on the mucosal microbiota and CDI severity. To
evaluate the mucosal microbiota, we sequenced snips of tissue collected
from the cecum, proximal colon, and distal colon. Similar to what was
observed with the stool samples, alpha diversity was lower in the
PEG-treated mice compared to clindamycin treatment (Fig. 3A). Although
alpha diversity continued to increase over time based on the communities
from PEG-treated mice collected at 20 and 30 days post-infection (Fig.
3A, Data Set S1, Sheet X). Group, time point, and their interactions
with other variables (cage, experiment number, and sample type)
explained the majority of the variation observed in mucosal communities
(PERMANOVA combined R\textsuperscript{2} = 0.83, \emph{P} \textless{}
0.05, Fig. 3B, Data Set S1, Sheet X). We saw the greatest difference in
the mucosal microbiota between treatment groups at 6 days post-infection
with 16 genera that were significantly different in all three of the
tissue types we collected (cecum, proximal colon, and distal colon; Fig.
S3A, Data Set S1, Sheet X). However, at 30 days post-infection only
\emph{Bacteroides}, \emph{Clostridiales}, \emph{Firmicutes}, and
\emph{Ruminococcaceae} were different between treatment groups and only
in the cecum tissues (Fig. 3C, Fig. 2E, Data Set S1, Sheet X).
Interestingly, one of the bacteria that was different between treatment
groups at 6 days post-infection was \emph{Peptostreptococcaceae} (the
genera that contains a sequence that matches \emph{C. difficile}).
\emph{Peptostreptococcaceae} was primarily only present in the 5-day PEG
treatment group of mice and decreased in the proximal and distal colon
tissues over time (Fig. S2B). Thus, PEG treatment had a significant
impact on the mucosal microbiota and we detected \emph{C. difficile}
sequences in the cecum, proximal colon, and distal colon tissue
communities.

@@

Previous work examining the impact of PEG on the murine microbiota found
that PEG treatment resulted in the permanent loss of S24-7, also known
as \emph{Muribaculum intestinale} (16). We decided to check our
\emph{Porphyromonadaceae} operational taxonomic units (OTUs) because
\emph{Muribaculum intestinale} is classified as
\emph{Porphyromonadaceae} by the Ribosomal Database Project (RDP)
database (31) and \emph{Porphyromonadaceae} was a top feature in the
random forest model predicting prolonged \emph{C. difficile}
colonization. We identified 4 OTUs that had at least 92\% identity to
\emph{Muribaculum intestinale} and examined their abundance in mice that
either cleared or were still colonized with \emph{C. difficile} at 10
days post-infection. While all of the OTUs, were decreased by PEG and
clindamycin treatment, there was some recovery in the mice that cleared
(Fig. S4A). We also examined other \emph{Porphyromonadaceae} and
\emph{Lachnospiraceae} OTUs since these were the 2 genera that were
important to our classification model and contained multiple OTUs that
were different at 5 days post-infection between mice that either cleared
or remained colonized with \emph{C. difficile} by 10 days post-infection
(Data Set S1, sheet X). While individual \emph{Porphyromonadaceae} and
\emph{Lachnospiraceae} OTUs tended to be more abundant in the mice that
clear \emph{C. difficile} relative to the mice that exhibit prolonged
colonization (Fig. S4B-C), there is no single OTU that fits the pattern
we observed at the genus level (Fig. 7C), suggesting multiple
\emph{Porphyromonadaceae} and \emph{Lachnospiraceae} OTUs influenced
\emph{C. difficile} clearance. Overall, our results suggest that
specific bacterial community differences explain the prolonged \emph{C.
difficile} colonization we observed in 5-day PEG and post-CDI 1-day PEG
treated mice.

@@

\textbf{Figure 1. 5-day PEG treatment prolongs susceptibility and mice
become persistently colonized with \emph{C. difficile}.} A. Setup of the
experimental time line for experiments with 5-day PEG treated mice
consisting of 4 treatment groups. 1. Clindamycin was administered at 10
mg/kg by intraperitoneal injection. 2. 15\% PEG 3350 was administered in
the drinking water for five days. 3. 5-day PEG plus clindamycin
treatment. 4. 5-day PEG plus 10-day recovery treatment. All treatment
groups were then challenged with 10\textsuperscript{5} \emph{C.
difficile} 630 spores. A subset of mice were euthanized on either 4 or 6
days post-infection and tissues were collected for histopathology
analysis, the remaining mice were followed through 20 or 30 days
post-infection. B. Weight change from baseline weight in groups after
treatment with PEG and/or clindamycin, followed by \emph{C. difficile}
challenge. C. \emph{C. difficile} CFU/gram stool measured over time (N =
16-59 mice per time point) via serial dilutions. The black line
represents the limit of detection for the first serial dilution. CFU
quantification data was not available for each mouse due to stool
sampling difficulties (particularly the day the mice came off of the PEG
treatment) or early deaths. For B-C, lines represent the median for each
treatment group and circles represent samples from individual mice.
Asterisks indicate time points where the weight change or CFU/g was
significantly different between groups by the Kruskal-Wallis test with
Benjamini-Hochberg correction for testing multiple time points. The data
presented are from a total of 5 separate experiments. \newpage

\includegraphics{figure_2.pdf} \textbf{Figure 2. 5-day PEG treatment
disrupts the stool microbiota for a longer amount of time compared to
clindamycin-treated mice.} A. Principal Coordinate analysis (PCoA) of
Bray-Curtis distances from stool samples collected throughout the
experiment. For A and C, each circle represents a sample from an
individual mouse and the transparency of the symbol corresponds to the
day post-infection. B. Bray-Curtis distances of stool samples collected
on either day 10 or 30 post-infection relative to the baseline sample
collected for each mouse (before any drug treatments were administered).
C. Shannon diversity in stool communities over time. The line indicates
the median value for each treatment group. The colors of the symbols and
lines represent the four treatment groups. D. 14 of the 33 genera
affected by PEG treatment (Data Set S1, sheet X). The symbols represent
the median relative abundance for a treatment group at either baseline
(open circle) or 1-day post treatment (closed circle). Data from the
5-day PEG and 5-day PEG plus 10-day recovery groups were analyzed by
paired Wilcoxan signed-rank test with Benjamini-Hochberg correction for
testing all identified genera. The clindamycin and 5-day PEG plus
clindamycin treatment groups are shown for comparison. E. 6 of the 24
genera that were significantly different between the four treatment
groups over multiple time points. Differences between treatment groups
were identified by Kruskal-Wallis test with Benjamini-Hochberg
correction for testing all identified genera. The gray vertical line (D)
and horizontal vertical lines (E) indicate the limit of detection.
\newpage

\includegraphics{figure_3.pdf} \textbf{Figure 3. 5-day PEG treatment
does not result in more severe CDIs, although mucosal microbiota is
altered.} A. Shannon diversity in cecum communities over time. The line
indicates the median value for each treatment group. The colors of the
symbols and lines represent the four treatment groups. A similar pattern
was observed with the proximal and distal colon communities (Data Set
S1, sheet X-X). B. PCoA of Bray-Curtis distances from mucosal samples
collected throughout the experiment. Circles, triangles, and squares
indicate cecum, proximal colon, and distal colon communities,
respectively. For A-B, transparency of the symbol corresponds to the day
post-infection that the sample was collected. C. The median relative
abundance of the 4 genera that were significantly different between the
cecum communities of different treatment groups on day 6 and day 30
(Data Set S1, sheet X). The gray vertical line indicate the limit of
detection. D-E. The histopathology summary scores from cecum and colon
H\&E stained slides. The summary score is the total score based on
evaluation of edema, cellular infiltration, and inflammation. Each
category is given a score ranging from 0-4, thus the maximum possible
summary score is 12. The tissue for histology was collected at either 4
(D) or 6 (E) days post-infection with the exception that one set of
5-day PEG treated mock-challenged mice were collected on day 0
post-infection (first set of open circles in D). Histology data were
analyzed with the Kruskal-Wallis test followed by pairwise Wilcoxon
comparisons with Benjamini-Hochberg correction. \newpage

\includegraphics{figure_4.pdf} \textbf{Figure 4. 1-day PEG treatment
renders mice susceptible to transient \emph{C. difficile} colonization.}
A. Setup of the experimental time line for the 1-day PEG treated mice
consisting of 3 treatment groups. 1. Clindamycin was administered at 10
mg/kg by intraperitoneal injection. 2. 15\% PEG 3350 was administered in
the drinking water for 1 day. 3. 1-day PEG plus 1-day recovery. The
three treatment groups were then challenged with 10\textsuperscript{5}
\emph{C. difficile} 630 spores. B. \emph{C. difficile} CFU/gram stool
measured over time (N = 12-18 mice per time point) via several
dilutions. The black dotted line represents the limit of detection for
the first serial dilution. Asterisks indicate time points where the
CFU/gram was significantly different between treatment groups by
Kruskall-Wallis test with Benjamini-Hochberg correction for testing
multiple time points. C. PCoA of Bray-Curtis distances from stool
communities collected from the three treatment groups over time (day:
R\textsuperscript{2} = 0.43; group: R\textsuperscript{2} = 0.19). D.
Shannon diversity in stool communities over time with colored lines
representing the median value for each treatment group. For B-D, each
symbol represents a sample from an individual mouse and symbol
transparency corresponds to the day post-infection that the sample was
collected. E. Median relative abundances per treatment group for 6 out
of the 14 genera that were affected by treatment, but recovered close to
baseline levels by 7 days post-infection (Fig. 3E, Data Set S1, Sheet
X). Stool samples from either baseline and day 1 or baseline and day 7
were analyzed by paired Wilcoxan signed-rank test with
Benjamini-Hochberg correction for testing all identified genera. The
gray horizontal line represents the limit of detection.

\includegraphics{figure_5.pdf} \textbf{Figure 5. 1-day PEG treatment
post \emph{C. difficile} challenge prolongs colonization regardless of
whether an FMT is also administered.} A. Setup off the experimental time
line for experiments with post-CDI PEG treated mice. There were a total
of 4 different treatment groups. All mice were administered 10 mg/kg
clindamycin intraperitoneally (IP) 1 day before challenge with
10\textsuperscript{3-5} \emph{C. difficile} 630 spores. 1. Did not
receive any additional treatment. 2. Immediately after \emph{C.
difficile} challenge, mice received 15\% PEG 3350 in the drinking water
for 1 day. 3-4. 3-days after challenge, mice received 1-day PEG
treatment and then received either a fecal microbiota transplant (3) or
PBS (4) solution by oral gavage. Mice were followed through 15-30 days
post-infection (only the post-CDI 1-day PEG group was followed through
30 days post-infection). B. CFU/g of \emph{C. difficile} stool measured
over time via serial dilutions. The black line represents the limit of
detection for the first serial dilution. C-D. Shannon diversity (C) and
richness (D) in stool communities over time. B-D. Each symbol represents
a stool sample from an individual mouse with the lines representing the
median value for each treatment group. The transparancy of the symbol
corresponds to the day post-infection. Asterisks indicate time points
with significant differences between groups by a Kruskall-Wallis test
with a Benjamini-Hochberg correction for testing multiple times points.
Colored rectangles indicates the 1-day PEG treatment period for
applicable groups. The data presented are from a total of 3 separate
experiments. \newpage

\includegraphics{figure_6.pdf} \textbf{Figure 6. For 1-day PEG treatment
post \emph{C. difficile} challenge mice that also receive an FMT only
some bacterial genera were restored.} A. PCoA of Bray-Curtis distances
from stool samples collected over time as well as the FMT solution that
was administered to one treatment group. Each circle represents an
individual sample, the transparency of the circle corresponds to day
post-infection as shown in Fig. 6C-D. B. Median relative abundances of 2
genera that were significantly different over multiple time points in
mice that were administered either FMT or PBS solution via gavage C.
Median relative abundances of the top 6 out of 24 genera that were
significant over multiple timepoints, plotted over time (Data Set S1,
Sheet X). For B-C, colored rectangles indicates the 1-day PEG treatment
period for applicable groups. Gray horizontal lines represent the limit
of detection.

@@

\includegraphics{figure_7.pdf} \textbf{Figure 7. Specific microbiota
features associated with prolonged \emph{C. difficile} colonization in
PEG treated mice.} A. Top ten bacteria that contributed to the random
forest model trained on five day post-infection community relative
abundance data to predict wehther mice would still be colonized with
\emph{C. difficile} 10 days post-infection. The median (point) and
interquantile range (lines) change in AUROC when the bacteria is left
out of the model is shown. B. The median relative abundances at 5 days
post-infection of the top ten bacteria that contributed to the random
forest classification model. Color indicates whether the mice were still
colonized with \emph{C. difficile} 10 days post-infection and the black
horizontal line represents the median relative abundance. Each symbol
represents a stool sample from an individual mouse and the shape of the
symbol indicates whether the PEG-treated mice recieved a 5-day (Fig.
1-3), 1-day (Fig. 4) or post-CDI PEG (Fig. 5-6) treatment. C. The median
relative abundances of the 5 genera with greater than 1\% median
relative abundance in the stool community over time. For B-C, the gray
horizontal line represents the limit of detection. \newpage

\includegraphics{figure_8.pdf} \textbf{Figure 8. Schematic summarizing
findings.} A. The gut microbiota of our C57Bl/6 mice is restant to
\emph{C. difficle} but treatment with either the antibiotic,
clindamycin, or the osmotic laxative, PEG 3350 renders the mice
susceptible to \emph{C. difficile} colonization. Recovery of
colonization resistance in clindamycin-treated mice is relatively
straightforward and the mice clear \emph{C.difficile} within 10 days
post-infection. However, for mice that recieved a 5-day PEG treatment or
a 1-day PEG treatment after \emph{C. difficile} challenge recovery of
colonization resistance is more uncertain because mice were still
colonized with \emph{C. difficile} 30 days post-infection in the case of
several PEG treatments. We found increased \emph{Porphyromonadaceae} and
\emph{Lachnospiraceae} were associated with recovery of colonization
resistance, while increased \emph{Enterobacteriaceae} and
\emph{Bacteroides} were associated with prolonged \emph{C. difficle}
colonization. B.Considering that most hospitals recommend not performing
\emph{C. difficile} testing on patients taking laxatives and laxatives
are used when administering fecal microbiota transplants via colonoscopy
to patients with recurrent CDIs. further studies are needed to evaluate
if laxatives impact human microbiota colonization resistance. Further
studies are needed to understand the impact of osmotic laxatives on
\emph{C. difficile} colonization resistance and clearance in human
patients. \newpage

\includegraphics{figure_S1.pdf} \textbf{Figure S1. Microbiota dynamics
post-infection in the 5-day PEG treatment plus 10-day recovery mice.} A.
\emph{C. difficile} CFU/g over time in the stool samples collected from
5-day PEG treatment plus 10-day recovery mice. Same data presented in
Fig. 1C, but the data for the other 3 treatment groups have been
removed. B. Median relative abundances of 8 bacterial genera from day 0
post-infection onward from the 10-day recovery mice. We analyzed samples
from day 0 and day 8 post-infection, which represented the the time
points where mice were challenged with \emph{C. difficile} and when the
median relative \emph{C. difficile} CFU stabilized for the group using
the paired Wilcoxan signed-rank test, but no genera were signifcant
after Benjamini-Hochberg correction. \newpage

\includegraphics{figure_S2.pdf} \textbf{Figure S2. PEG treatment still
has a large impact on the mucosal microbiota 6 days post-infection} A.
The\\
\newpage

@@

\includegraphics{figure_S4.pdf} \textbf{Figure S4. Specific OTUs
associated with clearance by 10 days post-infection that are mostly
absent in mice with prolonged \emph{C. difficile} colonization. Ex.
\emph{Muribaculum intestinale}.} A. \newpage
